---
title: "compare_peaks"
author: "caitlinpage"
date: "2024-09-04"
output: workflowr::wflow_html
editor_options:
  chunk_output_type: inline
---

## Introduction

```{r}
library(Damsel)
library(plyranges)
library(dplyr)
library(tidyr)
library(ggplot2)

library(BSgenome.Dmelanogaster.UCSC.dm6)
library(TxDb.Dmelanogaster.UCSC.dm6.ensGene)
library(org.Dm.eg.db)
library(ggVennDiagram)
library(rtracklayer)
```

# Compare Damsel and Vissers
```{r}
damsel_peaks <- readRDS("../output/damsel_peaks.rds")
vissers_peaks <- readRDS("../output/vissers_peaks.rds")
vissers_peaks_mod <- readRDS("../output/vissers_peaks_mod.rds")
```
```{r}
nrow(damsel_peaks)
nrow(vissers_peaks)
```

* Visser's peaks had to be modified in order to be usable (missing chromosome name)
* They also do not contain a p value - making it difficult to rank peaks and identify the most significant
```{r}
vissers_peaks_mod$peak_num <- 1:nrow(vissers_peaks_mod)
vissers_peaks_mod$Position <- paste0(vissers_peaks_mod$seqnames, "-", vissers_peaks_mod$start)
```


```{r}
peak_all <- union_ranges(as_granges(damsel_peaks), as_granges(vissers_peaks_mod)) %>% 
  data.frame() %>% mutate(compare_num = 1:n())
peak_compare_damsel <- find_overlaps(as_granges(peak_all), as_granges(damsel_peaks)) %>% data.frame()
peak_compare_vissers <- find_overlaps(as_granges(peak_all), as_granges(vissers_peaks_mod)) %>% data.frame()
```

```{r}
venn_data <- list(Damsel = peak_compare_damsel$compare_num, Vissers = peak_compare_vissers$compare_num)
ggVennDiagram(venn_data)
```

# Compare Damsel and Marshall
```{r}
marshall_peaks_1 <- rtracklayer::import("../data/sd_1_SRR794884-vs-Dam.kde-norm.gatc-FDR0.01.peaks.gff")
marshall_peaks_2 <- rtracklayer::import("../data/sd_2_SRR7948877-vs-Dam.kde-norm.gatc-FDR0.01.peaks.gff")
marshall_peaks <- readRDS("../output/marshall_peaks.rds")
```
```{r}
head(marshall_peaks_1)
```

```{r}
head(marshall_peaks_2)
```

* Marshall peaks had to be compiled in order to be analysed
* see `code/compile-Marshall_peaks.R`
```{r}
marshall_peaks
```


```{r}
peak_all <- union_ranges(as_granges(damsel_peaks), as_granges(marshall_peaks)) %>% 
  data.frame() %>% mutate(compare_num = 1:n())
peak_compare_damsel <- find_overlaps(as_granges(peak_all), as_granges(damsel_peaks)) %>% data.frame()
peak_compare_marshall <- find_overlaps(as_granges(peak_all), as_granges(marshall_peaks)) %>% data.frame()
```

```{r}
venn_data <- list(Damsel = peak_compare_damsel$compare_num, Marshall = peak_compare_marshall$compare_num)
ggVennDiagram(venn_data)
```

# Compare all 3 approaches
```{r}
peak_all <- union_ranges(union_ranges(as_granges(damsel_peaks), as_granges(vissers_peaks_mod)), as_granges(marshall_peaks)) %>% data.frame() %>% 
  mutate(compare_num = 1:n()) %>% as_granges() 

peak_compare_damsel <- find_overlaps(as_granges(peak_all), as_granges(damsel_peaks), maxgap = 150) %>% data.frame()
peak_compare_vissers <- find_overlaps(as_granges(peak_all), as_granges(vissers_peaks_mod), maxgap = 150) %>% data.frame()
peak_compare_marshall <- find_overlaps(as_granges(peak_all), as_granges(marshall_peaks), maxgap = 150) %>% data.frame()


venn_data <- list(Damsel = peak_compare_damsel$compare_num, Vissers = peak_compare_vissers$compare_num, 
                  Marshall = peak_compare_marshall$compare_num)
ggVennDiagram(venn_data)
```
```{r}
library(export)
```

```{r}
x11(width = dev.size("in")[[1]], height = dev.size("in")[[1]])
venn_plot <- ggVennDiagram(venn_data)
print(venn_plot)
graph2png(file="../outputvenn_plot.png",
         width = dev.size(units="px")[[1]]/40, 
         height = dev.size(units="px")[[1]]/40, 
         dpi = 900)
dev.off()
```

* what is the overlap with just Damsel's top 1000 peaks?
* peaks are ranked by FDR

```{r}
peak_all <- union_ranges(union_ranges(as_granges(damsel_peaks[1:1000,]), as_granges(vissers_peaks_mod)), as_granges(marshall_peaks)) %>% data.frame() %>% 
  mutate(compare_num = 1:n()) %>% as_granges() 

peak_compare_damsel <- find_overlaps(as_granges(peak_all), as_granges(damsel_peaks[1:1000,]), maxgap = 150) %>% data.frame()
peak_compare_vissers <- find_overlaps(as_granges(peak_all), as_granges(vissers_peaks_mod), maxgap = 150) %>% data.frame()
peak_compare_marshall <- find_overlaps(as_granges(peak_all), as_granges(marshall_peaks), maxgap = 150) %>% data.frame()


venn_data <- list(Damsel = peak_compare_damsel$compare_num, Vissers = peak_compare_vissers$compare_num, 
                  Marshall = peak_compare_marshall$compare_num)
ggVennDiagram(venn_data)
```
* overlap still looks pretty good
* Vissers' and Marshall are more similar to each other as they don't group the samples like Damsel does

* try with top 1000 Vissers' - ranked by logFC which is not ideal but just testing
```{r}
peak_all <- union_ranges(union_ranges(as_granges(damsel_peaks[1:1000,]), as_granges(vissers_peaks_mod[1:1000,])), as_granges(marshall_peaks)) %>% data.frame() %>% 
  mutate(compare_num = 1:n()) %>% as_granges() 

peak_compare_damsel <- find_overlaps(as_granges(peak_all), as_granges(damsel_peaks[1:1000,]), maxgap = 150) %>% data.frame()
peak_compare_vissers <- find_overlaps(as_granges(peak_all), as_granges(vissers_peaks_mod[1:1000,]), maxgap = 150) %>% data.frame()
peak_compare_marshall <- find_overlaps(as_granges(peak_all), as_granges(marshall_peaks), maxgap = 150) %>% data.frame()


venn_data <- list(Damsel = peak_compare_damsel$compare_num, Vissers = peak_compare_vissers$compare_num, 
                  Marshall = peak_compare_marshall$compare_num)
ggVennDiagram(venn_data)
```
* interesting that now Damsel has a slightly better overlap to Marshall than Vissers does
```{r}
peak_all <- union_ranges(union_ranges(as_granges(damsel_peaks[1:1000,]), as_granges(vissers_peaks_mod[1:1000,])), as_granges(marshall_peaks[1:1000,])) %>% data.frame() %>% 
  mutate(compare_num = 1:n()) %>% as_granges() 

peak_compare_damsel <- find_overlaps(as_granges(peak_all), as_granges(damsel_peaks[1:1000,]), maxgap = 150) %>% data.frame()
peak_compare_vissers <- find_overlaps(as_granges(peak_all), as_granges(vissers_peaks_mod[1:1000,]), maxgap = 150) %>% data.frame()
peak_compare_marshall <- find_overlaps(as_granges(peak_all), as_granges(marshall_peaks[1:1000,]), maxgap = 150) %>% data.frame()


venn_data <- list(Damsel = peak_compare_damsel$compare_num, Vissers = peak_compare_vissers$compare_num, 
                  Marshall = peak_compare_marshall$compare_num)
ggVennDiagram(venn_data)
```

```{r}
peak_all <- union_ranges(union_ranges(as_granges(damsel_peaks[1:1000,]), as_granges(vissers_peaks_mod)), as_granges(marshall_peaks[1:1000,])) %>% data.frame() %>% 
  mutate(compare_num = 1:n()) %>% as_granges() 

peak_compare_damsel <- find_overlaps(as_granges(peak_all), as_granges(damsel_peaks[1:1000,]), maxgap = 150) %>% data.frame()
peak_compare_vissers <- find_overlaps(as_granges(peak_all), as_granges(vissers_peaks_mod), maxgap = 150) %>% data.frame()
peak_compare_marshall <- find_overlaps(as_granges(peak_all), as_granges(marshall_peaks[1:1000,]), maxgap = 150) %>% data.frame()


venn_data <- list(Damsel = peak_compare_damsel$compare_num, Vissers = peak_compare_vissers$compare_num, 
                  Marshall = peak_compare_marshall$compare_num)
ggVennDiagram(venn_data)
```











